Cognate beginnings to bilingual lexical acquisition

### 2023-07-23 manuscript/_appendix/_syllable-frequency.qmd
Simplify dependencies

pool_tmp <- bvq_data$pool |>
    # drop items with missing observations in these variables
    drop_na(xsampa, wordbank_lemma) |>
    filter(n_lemmas == 1,
           # exclude items with more than two lemmas
           !is_multiword,
           # exclude multi-word items
           include,
           # exclude problematic items (e.g., multi-word items)
           te %in% duplicated_te) |>
    add_count(te, name = "n_te") |> # get only items with one translation in each language
    filter(n_te == 2) |>
    distinct(language, te, .keep_all = TRUE) |> 
    select(language, te, item, label, childes_lemma, xsampa) |> 
    mutate(syll = strsplit(xsampa, 
                           split = paste0(.sep = c("\\.", "\\\""), 
                                          collapse = "|")),
           syll = lapply(syll, function(x) x[x != ""]),
           n_syll = map_int(syll, length),
           item = str_remove(item, "cat_|spa_"))

items <- items |> 
    mutate(xsampa = ipa::ipa(ipa)) |> 
    mutate(syll = strsplit(xsampa, 
                           split = paste0(.sep = c("\\.", "\\\""), 
                                          collapse = "|")),
           syll = lapply(syll, function(x) x[x != ""]),
           n_syll = map_int(syll, length),
           item = str_remove(item, "cat_|spa_"))
```

@@

```{r fig-syll-freq}
#| label: fig-syll-freq
#| fig-cap: "Distribution of apositional syllable frequencies in Spanish and Catalan"
#| fig-height: 3.5
#| fig-width: 7
#| echo: false
#| warning: false
#| message: false
syll <- items |> 
    left_join(distinct(pool_tmp, te, childes_lemma),
              by = join_by(te)) |> 
    left_join(select(childes, token, freq_lex = freq_million),
              by = c("childes_lemma" = "token")) |> 
    select(language, item, te, meaning, xsampa, lv, freq_lex, syll) |> 
    unnest_longer(syll) |> 
    group_by(language, syll) |> 
    summarise(freq_syll = sum(freq_lex),
              .groups = "drop") |> 
    arrange(language, desc(freq_syll))

@@

```{r syllables-posterior}
#| label: syllables-posterior
#| echo: false
#| warning: false
#| message: false
rope_interval <- c(lower = -0.5, upper = 0.5)
# tidy predictor names
str_repl <- c(
    "b_Intercept" = "Intercept",
    "b_n_syll_std" = glue::glue("Syllables (+1 SD, {round(sd(syllables_data$n_syll), 3)})"),
    "b_lv_std" = glue::glue("Cognateness (+1 SD, {round(sd(syllables_data$lv), 2)})")
)

post_syllables <- model_fit_syllables |>
    bayestestR::describe_posterior(test = "rope",
                                   rope_range = rope_interval,
                                   ci_method = "HDI") |>
    as_tibble() |>
    janitor::clean_names() |>
    mutate(.variable_name = factor(
        parameter,
        levels = names(str_repl),
        labels = str_repl) |>
            as.character(),
        parameter = ifelse(grepl("Intercept", parameter),
                           stringr::str_remove_all(parameter, "Intercept \\(|\\)"),
                           parameter)) |>
    select(.variable = parameter,
           .variable_name,
           .median = median,
           .upper = ci_high,
           .lower = ci_low,
           .rope_overlap = rope_percentage) |>
    ungroup()

post_draws_list <- post_syllables |> 
    split(post_syllables$.variable)

@@

We fit this model running `r dim(model_fit_syllables$fit)[2]` sampling chains with `r format(dim(model_fit_syllables$fit)[1], big.mark = ",")` iterations each. @tbl-syll-coefs shows a summary of the posterior distribution of the fixed effects in the model. As expected, words with more syllables scored higher in syllabic frequency: all posterior draws for the regression coefficient of the main effect of this predictor fell outside the ROPE defined between -0.5 and +0.5 ($\beta$ = `r round(post_draws_list[["b_n_syll_std"]]$.median, 2)`, 95% HDI = [`r round(post_draws_list[["b_n_syll_std"]]$.lower, 2)`, `r round(post_draws_list[["b_n_syll_std"]]$.upper, 2)`]). Keeping the number of syllables constant, the effect of cognateness was negligible: all of the posterior distributions of this predictor fell within the ROPE, providing evidence that the true value of the increment in syllabic frequency for every increase in cognateness is equivalent to zero ($\beta$ = `r round(post_draws_list[["b_lv_std"]]$.median, 2)`, 95% HDI = [`r round(post_draws_list[["b_lv_std"]]$.lower, 2)`, `r round(post_draws_list[["b_lv_std"]]$.upper, 2)`]).

```{r tbl-syll-coefs}
#| label: tbl-syll-coefs
#| tbl-cap: "Posterior distribution of regression coefficients."
#| tbl-note: "Median: median of the posterior distribution in the probability scale. 95\\% HDI: 95\\% highest density interval of the distribution; narrowest range of values that contains 95\\% of the distribution, and therefore is 95\\% likely to contain the true value, given the data. On top of each distribution, we indicate the proportion of posterior samples in the 95\\% HDI that fall into the region of practical equivalence (ROPE). This proportion indicates the probability of the true value of the coefficient being equivalent to zero. Lower values indicate that the true value is unlikely to be zero of equivalent to zero."
#| echo: false
#| message: false
#| warning: false
# summarise posterior draws
tbl_data <- post_syllables |>
    select(.variable_name, .median, .lower, .upper, .rope_overlap) |>
    mutate(
        .variable_name = gsub("×", '$\\\\times$', .variable_name),
        .hdi = paste0("[", round(.lower, 3),
                      ", ", round(.upper, 3), "]"),
        .rope_overlap = sub("^(-?)0.", "\\1.",
                            sprintf("%.3f", .rope_overlap))) |> 
    select(.variable_name, .median, .lower, .upper, .hdi, .rope_overlap)

tbl_data |> 
    mutate(viz = "") |> 
    select(-c(.lower, .upper)) |> 
    knitr::kable(
        format = "latex", 
        escape = FALSE,
        col.names = c("",
                      "$\\beta$",
                      "95\\% HDI",
                      "$p(\\text{ROPE})$",
                      ""),
        digits = 3,
        booktabs = TRUE,
        align = "lclrc",
        linesep = ""
    ) |> 
    kableExtra::kable_styling() |> 
    kableExtra::column_spec(
        5,
        background = "#f2f2f2",
        image = kableExtra::spec_pointrange(
            x = tbl_data$.median, 
            xmin = tbl_data$.lower, 
            xmax = tbl_data$.upper, 
            vline = 0,
            xaxt = "n",
            col = "#000000",
            res = 200)
    )
```

@@

```{r fig-syll-marginal}
#| label: fig-syll-marginal
#| fig-cap: "Posterior-predictions of the syllabic frequency model. Thicker lines indicate the median of the posterior predictions, and thinner lines indicate individual posterior predictions."
#| fig-width: 7
#| fig-height: 5
#| out-width: 80%
#| echo: false
#| message: false
#| warning: false
expand_grid(lv_std = scale(seq(0, 1, 0.1),
                           center = mean(syllables_data$lv),
                           scale = sd(syllables_data$lv))[, 1],
            n_syll_std = scale(seq(1, 4, 1),
                               center = mean(syllables_data$n_syll),
                               scale = sd(syllables_data$n_syll))[, 1]) |> 
    add_epred_draws(model_fit_syllables,
                    re_formula = NA) |>
    median_hdci() |> 
    mutate(
        lv_std = (lv_std * sd(syllables_data$lv)) + mean(syllables_data$lv),
        n_syll_std = (n_syll_std * sd(syllables_data$n_syll)) + mean(syllables_data$n_syll),
        n_syll_std = as.factor(n_syll_std)) |> 
    ggplot(aes(lv_std, .epred, 
               shape = n_syll_std)) +
    geom_point(data = filter(syllables_data, n_syll < 5) |> 
                   mutate(n_syll_std = as.factor(n_syll)),
               aes(x = lv,
                   y = freq_syll,
                   shape = n_syll_std),
               alpha = 1/2) +
    # geom_ribbon(aes(ymin = .lower,
    #                 ymax = .upper),
    #             alpha = 1/2,
    #             colour = NA,
    #             show.legend = FALSE) +
    geom_line(aes(linetype = n_syll_std),
              linewidth = 0.5) +
    labs(x = "Cognateness (Levenshtein similarity)",
         y = "Syllabic frequency",
         linetype = "Number of syllables",
         shape = "Number of syllables",
         colour = "Number of syllables") +
    scale_y_continuous(n.breaks = 8) +
    theme(panel.grid = element_blank(),
          legend.position = "top")
```

>>>>

pool_tmp <- bvq_data$pool |>
    # drop items with missing observations in these variables
    drop_na(xsampa, wordbank_lemma) |>
    filter(n_lemmas == 1,
           # exclude items with more than two lemmas
           !is_multiword,
           # exclude multi-word items
           include,
           # exclude problematic items (e.g., multi-word items)
           te %in% duplicated_te) |>
    add_count(te, name = "n_te") |> # get only items with one translation in each language
    filter(n_te == 2) |>
    distinct(language, te, .keep_all = TRUE) |> 
    select(language, te, item, label, childes_lemma, xsampa) |> 
    mutate(syll = strsplit(xsampa, 
                           split = paste0(.sep = c("\\.", "\\\""), 
                                          collapse = "|")),
           syll = lapply(syll, function(x) x[x != ""]),
           n_syll = purrr::map_int(syll, length),
           item = gsub("cat_|spa_", "", item))

items <- items |> 
    mutate(xsampa = ipa::ipa(ipa)) |> 
    mutate(syll = strsplit(xsampa, 
                           split = paste0(.sep = c("\\.", "\\\""), 
                                          collapse = "|")),
           syll = lapply(syll, function(x) x[x != ""]),
           n_syll = purrr::map_int(syll, length),
           item = gsub("cat_|spa_", "", item)) 
```

@@

```{r fig-syll-freq}
#| label: fig-syll-freq
#| fig-cap: "Distribution of apositional syllable frequencies in Spanish and Catalan"
#| fig-height: 3.5
#| fig-width: 7
#| echo: false
#| warning: false
#| message: false
syll <- items |> 
    left_join(distinct(pool_tmp, te, childes_lemma),
              by = join_by(te)) |> 
    left_join(select(childes, token, freq_lex = freq_million),
              by = c("childes_lemma" = "token")) |> 
    select(language, item, te, meaning, xsampa, lv, freq_lex, syll) |> 
    unnest_longer(syll) |> 
    summarise(freq_syll = sum(freq_lex),
              .by = c(language, syll)) |> 
    arrange(language, desc(freq_syll))

@@

```{r syllables-posterior}
#| label: syllables-posterior
#| echo: false
#| warning: false
#| message: false
rope_interval <- c(lower = -0.5, upper = 0.5)
# tidy predictor names

post_draws_list <- posterior_syllables_summary |> 
    split(posterior_syllables_summary$.variable)

@@

We fit this model running `r dim(model_fit_syllables$fit)[2]` sampling chains with `r format(dim(model_fit_syllables$fit)[1], big.mark = ",")` iterations each. @tbl-syll-coefs shows a summary of the posterior distribution of the fixed effects in the model. As expected, words with more syllables scored higher in syllabic frequency: all posterior draws for the regression coefficient of the main effect of this predictor fell outside the ROPE defined between -0.5 and +0.5 ($\beta$ = `r round(post_draws_list[["b_n_syll_std"]]$.value, 2)`, 95% HDI = [`r round(post_draws_list[["b_n_syll_std"]]$.lower, 2)`, `r round(post_draws_list[["b_n_syll_std"]]$.upper, 2)`]). Keeping the number of syllables constant, the effect of cognateness was negligible: all of the posterior distributions of this predictor fell within the ROPE, providing evidence that the true value of the increment in syllabic frequency for every increase in cognateness is equivalent to zero ($\beta$ = `r round(post_draws_list[["b_lv_std"]]$.value, 2)`, 95% HDI = [`r round(post_draws_list[["b_lv_std"]]$.lower, 2)`, `r round(post_draws_list[["b_lv_std"]]$.upper, 2)`]).

```{r tbl-syll-coefs}
#| label: tbl-syll-coefs
#| tbl-cap: "Posterior distribution of regression coefficients."
#| tbl-note: "Median: median of the posterior distribution in the probability scale. 95\\% HDI: 95\\% highest density interval of the distribution; narrowest range of values that contains 95\\% of the distribution, and therefore is 95\\% likely to contain the true value, given the data. On top of each distribution, we indicate the proportion of posterior samples in the 95\\% HDI that fall into the region of practical equivalence (ROPE). This proportion indicates the probability of the true value of the coefficient being equivalent to zero. Lower values indicate that the true value is unlikely to be zero of equivalent to zero."
#| echo: false
#| message: false
#| warning: false
# summarise posterior draws
tbl_data <- posterior_syllables_summary |>
    select(.variable_name, .value, .lower, .upper, .rope) |>
    mutate(
        .variable_name = gsub("×", '$\\\\times$', .variable_name),
        .hdi = paste0("[", round(.lower, 3),
                      ", ", round(.upper, 3), "]"),
        .rope = sub("^(-?)0.", "\\1.",
                    sprintf("%.3f", .rope))) |> 
    select(.variable_name, .value, .lower, .upper, .hdi, .rope)

tbl_data |> 
    mutate(viz = "") |> 
    select(-c(.lower, .upper)) |> 
    knitr::kable(
        format = "latex", 
        escape = FALSE,
        col.names = c("",
                      "$\\beta$",
                      "95\\% HDI",
                      "$p(\\text{ROPE})$"),
        digits = 3,
        booktabs = TRUE,
        align = "lclr",
        linesep = ""
    ) |> 
    kableExtra::kable_styling()
```

@@

```{r fig-syll-marginal}
#| label: fig-syll-marginal
#| fig-cap: "Posterior-predictions of the syllabic frequency model. Thicker lines indicate the median of the posterior predictions, and thinner lines indicate individual posterior predictions."
#| fig-width: 7
#| fig-height: 5
#| out-width: 80%
#| echo: false
#| message: false
#| warning: false
expand_grid(lv_std = scale(seq(0, 1, 0.1),
                           center = mean(syllables_data$lv),
                           scale = sd(syllables_data$lv))[, 1],
            n_syll_std = scale(seq(1, 4, 1),
                               center = mean(syllables_data$n_syll),
                               scale = sd(syllables_data$n_syll))[, 1]) |> 
    tidybayes::add_epred_draws(model_fit_syllables,
                               re_formula = NA) |>
    tidybayes::median_hdci() |> 
    mutate(
        lv_std = (lv_std * sd(syllables_data$lv)) + mean(syllables_data$lv),
        n_syll_std = (n_syll_std * sd(syllables_data$n_syll)) + mean(syllables_data$n_syll),
        n_syll_std = as.factor(n_syll_std)) |> 
    ggplot(aes(lv_std, .epred, 
               shape = n_syll_std)) +
    geom_point(data = filter(syllables_data, n_syll < 5) |> 
                   mutate(n_syll_std = as.factor(n_syll)),
               aes(x = lv,
                   y = freq_syll,
                   shape = n_syll_std),
               alpha = 1/2) +
    # geom_ribbon(aes(ymin = .lower,
    #                 ymax = .upper),
    #             alpha = 1/2,
    #             colour = NA,
    #             show.legend = FALSE) +
    geom_line(aes(linetype = n_syll_std),
              linewidth = 0.5) +
    labs(x = "Cognateness (Levenshtein similarity)",
         y = "Syllabic frequency",
         linetype = "Number of syllables",
         shape = "Number of syllables",
         colour = "Number of syllables") +
    scale_y_continuous(n.breaks = 8) +
    theme(panel.grid = element_blank(),
          legend.position = "top")
```

### 2023-07-23 manuscript/appendix.tex
Simplify dependencies

\usepackage{lineno}
\linenumbers
\newcommand{\runninghead}{A Preprint }
\renewcommand{\runninghead}{Cognate beginnings to bilingual lexical
acquistion (Appendix) }
\author{}
\date{}
\begin{document}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, breakable, sharp corners, boxrule=0pt, frame hidden]}{\end{tcolorbox}}\fi

@@

\begin{tabular}{lclrl}
\toprule
 & $\beta$ & 95\% HDI & $p(\text{ROPE})$ & \\
\midrule
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Model: Exposure}}\\
\hspace{1em}Age (+1 SD, 4.87, months) & 0.600 & {}[0.588, 0.611] & .000 & \\
\hspace{1em}Exposure (+1 SD, 1.81) & 0.558 & {}[0.55, 0.567] & .000 & \\
\hspace{1em}Cognateness (+1 SD, 0.26) & 0.515 & {}[0.504, 0.526] & .037 & \\
\hspace{1em}Length (+1 SD, 1.56 phonemes) & 0.485 & {}[0.478, 0.491] & .000 & \\
\hspace{1em}Age $\times$ Exposure & 0.518 & {}[0.51, 0.526] & .000 & \\
\hspace{1em}Age $\times$ Cognateness & 0.504 & {}[0.5, 0.506] & .985 & \\
\hspace{1em}Exposure $\times$ Cognateness & 0.486 & {}[0.483, 0.488] & .000 & \\
\hspace{1em}Age $\times$ Exposure $\times$ Cognateness & 0.495 & {}[0.493, 0.498] & .975 & \\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Model: Frequency \& DoE}}\\
\hspace{1em}Age (+1 SD, 4.87, months) & 0.601 & {}[0.59, 0.612] & .000 & \\
\hspace{1em}DoE (+1 SD, 0.3) & 0.558 & {}[0.549, 0.566] & .000 & \\
\hspace{1em}Cognateness (+1 SD, 0.26) & 0.516 & {}[0.506, 0.528] & .016 & \\
\hspace{1em}Phonemes (+1 SD, 1.56 phonemes) & 0.486 & {}[0.479, 0.492] & .000 & \\
\hspace{1em}Frequency (+1 SD, 0.19) & 0.527 & {}[0.516, 0.54] & .000 & \\
\hspace{1em}Age $\times$ DoE & 0.518 & {}[0.51, 0.527] & .000 & \\
\hspace{1em}Age $\times$ Cognateness & 0.504 & {}[0.501, 0.507] & .954 & \\
\hspace{1em}DoE $\times$ Cognateness & 0.486 & {}[0.483, 0.488] & .000 & \\
\hspace{1em}Age $\times$ DoE $\times$ Cognateness & 0.495 & {}[0.493, 0.498] & .962 & \\
\bottomrule
\end{tabular}

@@

\centering
\begin{tabular}{lllcllc}
\toprule
Translation & Item & X-SAMPA & Syllables & Item & X-SAMPA & Syllables\\
\midrule
ladybug & marieta & m@.4i"E.t@ & 4 & mariquita & ma.4i"ki.ta & 4\\
beans & mongetes & muJ"ZE.t@s & 3 & alubias & a"lu.bjas & 3\\
tongue & llengua & "LeN.gw@ & 2 & lengua & "len.gwa & 2\\
kitchen & cuina & "kuj.n@ & 2 & cocina & ko"Ti.na & 3\\
cheek & galta & "ga5.t@ & 2 & mejilla & me"xi.La & 3\\
\addlinespace
purse & bossa & "bo.s@ & 2 & bolso & "bol.so & 2\\
farm & granja & g4aJ.Z@ & 2 & granja & "g4an.xa & 2\\
sausage & salsitxa & s@5"si.t͡S@ & 3 & salchicha & sal"tSi.tSa & 3\\
cereal & cereals & s@.4e"a5s & 3 & cereales & Te.4e"a.les & 4\\
bottle & ampolla & @m"po.L@ & 3 & botella & bo"te.La & 3\\
\addlinespace
boat & vaixell & b@"SeL & 2 & barco & "ba4.ko & 2\\
family & família & fa"mi.lia & 3 & familia & fa"mi.lia & 3\\
balloon & globus & "g5O.Bus & 2 & globo & "glo.Bo & 2\\
bee & abella & @"BE.L@ & 3 & abeja & a"Be.xa & 3\\
earring & arracada / arracades & @.r@"ka.D@ & 4 & pendientes & pen"Djen.tes & 3\\
\bottomrule
\end{tabular}
\end{table}

@@

We fit this model running 4 sampling chains with 1,000 iterations each.
Table~\ref{tbl-syll-coefs} shows a summary of the posterior distribution
of the fixed effects in the model. As expected, words with more
syllables scored higher in syllabic frequency: all posterior draws for
the regression coefficient of the main effect of this predictor fell
outside the ROPE defined between -0.5 and +0.5 (\(\beta\) = 5.64, 95\%
HDI = {[}5.58, 5.71{]}). Keeping the number of syllables constant, the
effect of cognateness was negligible: all of the posterior distributions
of this predictor fell within the ROPE, providing evidence that the true
value of the increment in syllabic frequency for every increase in
cognateness is equivalent to zero (\(\beta\) = 0.01, 95\% HDI =
{[}-0.06, 0.07{]}).

@@

\centering
\begin{tabular}{lclr>{}c}
\toprule
 & $\beta$ & 95\% HDI & $p(\text{ROPE})$ & \\
\midrule
Intercept & 16.089 & {}[16.023, 16.163] & .000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/appendix_files/figure-latex/pointrange_1e8078876ee9.pdf}\\
Syllables (+1 SD, 0.802) & 5.642 & {}[5.577, 5.711] & .000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/appendix_files/figure-latex/pointrange_1e807c1b3ef.pdf}\\
Cognateness (+1 SD, 0.24) & 0.007 & {}[-0.059, 0.075] & 1.000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/appendix_files/figure-latex/pointrange_1e8043796412.pdf}\\
\bottomrule
\end{tabular}
\end{table}

>>>>

\usepackage{lineno}
\linenumbers
\newcommand{\runninghead}{A Preprint }
\renewcommand{\runninghead}{Cognate beginnings to bilingual lexical
acquistion (Appendix) }
\author{}
\date{}
\begin{document}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, frame hidden, sharp corners, breakable, enhanced, boxrule=0pt]}{\end{tcolorbox}}\fi

@@

\begin{tabular}{lclrl}
\toprule
 & $\beta$ & 95\% HDI & $p(\text{ROPE})$ & \\
\midrule
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Model: Exposure}}\\
\hspace{1em}Age (+1 SD, 4.86 months) & 0.600 & {}[0.588, 0.611] & .000 & \\
\hspace{1em}Age $\times$ Exposure & 0.518 & {}[0.51, 0.526] & .000 & \\
\hspace{1em}Age $\times$ Exposure $\times$ Cognateness & 0.495 & {}[0.493, 0.498] & .907 & \\
\hspace{1em}Age $\times$ Cognateness & 0.504 & {}[0.5, 0.506] & .974 & \\
\hspace{1em}Exposure (+1 SD, 1.81) & 0.558 & {}[0.55, 0.566] & .000 & \\
\hspace{1em}Exposure $\times$ Cognateness & 0.486 & {}[0.483, 0.489] & .000 & \\
\hspace{1em}Cognateness (+1 SD, 0.26) & 0.515 & {}[0.504, 0.527] & .089 & \\
\hspace{1em}Length (+1 SD, 1.56 phonemes) & 0.485 & {}[0.478, 0.491] & .000 & \\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Model: Frequency \& DoE}}\\
\hspace{1em}Age (+1 SD, 4.86, months) & 0.601 & {}[0.59, 0.612] & .000 & \\
\hspace{1em}Age $\times$ DoE & 0.518 & {}[0.51, 0.527] & .000 & \\
\hspace{1em}Age $\times$ DoE $\times$ Cognateness & 0.495 & {}[0.493, 0.498] & .901 & \\
\hspace{1em}Age $\times$ Cognateness & 0.504 & {}[0.501, 0.507] & .898 & \\
\hspace{1em}DoE (+1 SD, 0.3) & 0.558 & {}[0.549, 0.566] & .000 & \\
\hspace{1em}DoE $\times$ Cognateness & 0.486 & {}[0.483, 0.488] & .000 & \\
\hspace{1em}Frequency (+1 SD, 0.19) & 0.527 & {}[0.516, 0.54] & .000 & \\
\hspace{1em}Cognateness (+1 SD, 0.26) & 0.516 & {}[0.505, 0.527] & .043 & \\
\hspace{1em}Phonemes (+1 SD, 1.56 phonemes) & 0.486 & {}[0.479, 0.492] & .000 & \\
\bottomrule
\end{tabular}

@@

\centering
\begin{tabular}{lllcllc}
\toprule
Translation & Item & X-SAMPA & Syllables & Item & X-SAMPA & Syllables\\
\midrule
drink (beverage) & beguda & b@"Gu.D@ & 3 & bebida & be"Bi.Da & 3\\
snail & cargol & k@r"gOl & 2 & caracol & ka.4a"kol & 3\\
red & vermell & b@r"mEL & 2 & rojo & "ro.xo & 2\\
book & conte & "kon.t@ & 2 & cuento & "kwen.to & 2\\
hand & mà / mans & m@ & 1 & manos & "ma.nos & 2\\
\addlinespace
egg & ou & Ow & 1 & huevo & "we.Bo & 2\\
pool & piscina & pi"si.n@ & 3 & piscina & pis"Ti.na & 3\\
fox & guineu & gi"nEw & 2 & zorro & "To.ro & 2\\
potato & patata & p@"ta.t@ & 3 & patata & pa"ta.ta & 3\\
arm & braç / braços & b4as & 1 & brazos & "b4a.Tos & 2\\
\addlinespace
cow & vaca & "ba.k@ & 2 & vaca & "ba.ka & 2\\
sausage & salsitxa & s@5"si.t͡S@ & 3 & salchicha & sal"tSi.tSa & 3\\
tongue & llengua & "LeN.gw@ & 2 & lengua & "len.gwa & 2\\
park & parc & park & 1 & corralito & ko.ra"li.to & 4\\
black & negre & "nE.G4@ & 2 & negro & "ne.G4o & 2\\
\bottomrule
\end{tabular}
\end{table}

@@

We fit this model running 4 sampling chains with 1,000 iterations each.
Table~\ref{tbl-syll-coefs} shows a summary of the posterior distribution
of the fixed effects in the model. As expected, words with more
syllables scored higher in syllabic frequency: all posterior draws for
the regression coefficient of the main effect of this predictor fell
outside the ROPE defined between -0.5 and +0.5 (\(\beta\) = 5.64, 95\%
HDI = {[}5.57, 5.71{]}). Keeping the number of syllables constant, the
effect of cognateness was negligible: all of the posterior distributions
of this predictor fell within the ROPE, providing evidence that the true
value of the increment in syllabic frequency for every increase in
cognateness is equivalent to zero (\(\beta\) = 0.01, 95\% HDI =
{[}-0.06, 0.07{]}).

@@

\centering
\begin{tabular}{lclr>{}c}
\toprule
 & $\beta$ & 95\% HDI & $p(\text{ROPE})$ & \\
\midrule
Intercept & 16.089 & {}[16.022, 16.166] & .000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/gonza/Documents/cognate-beginnings/appendix_files/figure-latex/pointrange_b043454131.pdf}\\
Cognateness (+1 SD, 0.24) & 0.008 & {}[-0.057, 0.074] & 1.000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/gonza/Documents/cognate-beginnings/appendix_files/figure-latex/pointrange_b0479223d59.pdf}\\
Syllables (+1 SD, 0.802) & 5.642 & {}[5.574, 5.709] & .000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/gonza/Documents/cognate-beginnings/appendix_files/figure-latex/pointrange_b042ac2624.pdf}\\
\bottomrule
\end{tabular}
\end{table}
